Bayesian Inference Using Sequential Monte-Carlo Algorithm for Dynamic System Models

### 2020-07-22 MSc-dissertation.tex
New para

\usepackage{epcc}
\usepackage{graphics}
% \usepackage{mathtools}
\usepackage{amsmath}

@@

\chapter{Implementations and Experiments}

@@

@@

\begin{figure}
    \begin{center}
    \resizebox{1.0\hsize}{!}{\includegraphics{fig/LS.png}}
    \end{center}

@@

\subsection{Implementations and code}

@@

>>>>

\usepackage{epcc}
\usepackage{graphics}
\usepackage{hyperref}
% \usepackage{mathtools}
\usepackage{amsmath}

@@

\chapter{Parameter estimation and model comparison}

@@

\section{Implementations and code}

According to our preliminary studies, ABC SMC framework implemented using \verb|pyABC| in \verb|Python| is adopted in this project. \verb|pyABC| is a popular SMC software packages \cite{ref:pyabc} that has been used form many SMC studies and applications \cite{ref:adpt_pop}. \verb|pyABC| provides an open-source framework for likelihood-free inference, which is an implementation of ABC SMC algorithm and a tool-box for inference and analysis tasks. Besides, it supports multi-core execution implemented using \verb|multiprocessing| package for scaling-up, which is of our interests and suitable for the followed performance experiments.

The code for the project is developed and tested in local environment (macOS 10.15.6, see Table \ref{table:local_macine}) at first, then the functional version is deployed on compute node of ARCHER and Cirrus for parameter inference experiments and model selection task. The scaling-up experiment is performed on Cirrus, the profiling and its analysis are performed on local machine. Some development and debug are on the remote machines to ensure experiments run correctly on compute node.

The ODE model-related code and utilities are firstly developed to enable reusable functions, variables and data calls. It includes the code for the ODE models, ODE solver, data structures, data format transform functions etc. By doing this in the implementation code file we can set options and activate ABC SMC easily, without creating definition and deplicated code for the models and parameter priors.

Some modifications are made to the source code of \verb|pyABC|. For example, the laboratorial measured data are not considered `complete' for modeling and analysis purpose: there are different measurement time points for cells and cytokines, e.g. cell numbers is not measured at 0.25 hpl and cytokines' expression is not measured at time point 120 hpl. These missing values are treated as `ignored' when calculating the distance between observed data and simulated data. To cope with this built-in distance function code is modified.

Analysis code is also wrapped up into functions which can be easily called after an ABC SMC run is finished. They includes multiple visualisations and summary statistic of the results. For different tasks there are also separated analysis code file. Other analysis tools used in the project include built-in database visualisation tool, R code and Microsoft Excel.

The project code is managed via \verb|git| version-control and available as a GitHub repository\footnote{\url{https://github.com/chaolinhan/MSc_Project}}; it also make the cross machine synchronisation of code versions convenient.

After several development iterations, the code files for ABC SMC implementation now mainly has two types: code file for infer-back experiments (synthetic data), code for parameter inference (experimental data). We can just add more models in these code files for a model comparison task.

@@

\begin{figure}[h!]
    \begin{center}
    \resizebox{1.0\hsize}{!}{\includegraphics{fig/LS.png}}
    \end{center}

@@

@@

\subsection{Local machine}

Local development machine is a Mac laptop, running on macOS 10.15.6. The environment of the development is listed in Table \ref{table:local_macine}.

\begin{table}[h!]
    \centering
    \begin{tabular}{|c c|} 
     \hline
     Environment & Version \\ [0.5ex] 
     \hline\hline
     $\lambda_N$ & 2.1989  \\ 
     $\kappa_{N\beta}$ & 3.9627  \\
     $\mu_N$ & 1.7219 \\
     $\nu_{N\Phi}$  0.2195 & $cell^{-1}\cdotp h^{-1}$ \\
     \hline
     $\lambda_\Phi$ & 1.3146  $cell/h$ \\
     $\kappa_{\Phi\beta}$ 0.1235 & $cell/(unit\cdotp h)$ \\
     $\mu_\Phi$  0.1454 & $h^{-1}$ \\
     \hline
     $s_{\beta N}$ &6.5536  $unit/(cell\cdotp h)$ \\
     $i_{\beta\Phi}$  1.7062 & $cell^{-1}$ \\
     $\mu_\beta$ &0.5212  $h^{-1}$ \\
     \hline
     $s_{\alpha\Phi}$  10.2416 & $unit/(cell\cdotp h)$ \\
     $\mu_\alpha$ 19.6642 & $h^{-1}$ \\
    [1ex] 
     \hline
    \end{tabular}
    \caption{Environment on local machine}
    \label{table:local_macine}
\end{table}

\subsection{Remote machine}

\begin{table}[h!]
    \centering
    \begin{tabular}{|c c|} 
     \hline
     Environment & Version \\ [0.5ex] 
     \hline\hline
     $\lambda_N$ & 2.1989  \\ 
     $\kappa_{N\beta}$ & 3.9627  \\
     $\mu_N$ & 1.7219 \\
     $\nu_{N\Phi}$  0.2195 & $cell^{-1}\cdotp h^{-1}$ \\
     \hline
     $\lambda_\Phi$ & 1.3146  $cell/h$ \\
     $\kappa_{\Phi\beta}$ 0.1235 & $cell/(unit\cdotp h)$ \\
     $\mu_\Phi$  0.1454 & $h^{-1}$ \\
     \hline
     $s_{\beta N}$ &6.5536  $unit/(cell\cdotp h)$ \\
     $i_{\beta\Phi}$  1.7062 & $cell^{-1}$ \\
     $\mu_\beta$ &0.5212  $h^{-1}$ \\
     \hline
     $s_{\alpha\Phi}$  10.2416 & $unit/(cell\cdotp h)$ \\
     $\mu_\alpha$ 19.6642 & $h^{-1}$ \\
    [1ex] 
     \hline
    \end{tabular}
    \caption{Environment on remote machine}
    \label{table:remote_macine}
\end{table}
